---
title: "Prediction Assignment Writeup"
author: "Rafael Garcia Cunha"
date: "Saturday, April 25, 2015"
output: html_document
---

### Packages
```{r}
library(caret)
library(randomForest)
```

### Getting and Cleaning Data
```{r}
train <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""), h = T)
train <- train[,-1] #removing the ID row variable
test <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""), h = T)
test <- test[,-1] #removing the ID row variable
```

I read both data sets and removed the first colum of each cause it is an ID. I aso converted blanks ('""'), '#DIV/0' and 'NA' into NA values.

### Partitionating Data
```{r}
inTrain <- createDataPartition(train$classe, p=0.6, list=FALSE)
training <- train[inTrain, ]
testing <- train[-inTrain, ]
dim(training); dim(testing)
```

I'm dividing data to 60% training and 40% testing sets.

### Variable Selection
```{r}
hist(colSums(is.na(training)/dim(training)[1]), plot=F)
variables <- c(colSums(is.na(training)/dim(training)[1]) < 0.2)
training <- training[,variables]
testing <- testing[,variables]
```

As we can see from the \textbf{hist} function, we have 100 variables with more than `r 0.8` rate of NA. So we'll work with the other 59. 

### Model and Predictions

I'm fitting a random forest model with \textit{classe} as outcome and the other 58 as possible predictors. I'm also looking the variable importance measures produced by the random forest. 
```{r}
fitRF <- randomForest(classe ~ ., data = training)
varImp(fitRF)

pred1 <- predict(fitRF, newdata = training)
confusionMatrix(pred1, training$classe)

pred2 <- predict(fitRF, newdata = testing)
confusionMatrix(pred2, testing$classe)
```

The \textit{confusionMatrix} is a tool to evaluate the fitted model. In the training data we have Accuracy of 1, Sensitivity and Specificity all equal to 1 too. In the testing data the accuracy is 99.86% (so the expected out-of-sample error is 0.14%) and very good sensitivity and specificity.

### Test Data
```{r}
test <- test[,variables]
test <- test[,-ncol(test)]
```

```{r, echo = F}
test <- rbind(training[100, -59], test) 
row.names(test) <- c(100, 1:20)
test <- test[-1,]
```

```{r}
pred3 <- predict(fitRF, newdata = test)
pred3

setwd("C:\\Users\\Rafael\\Documents\\Coursera\\Practical Machine Learning")

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(pred3)
```